---
title: "Supplementary Information 2: Critical strength of interactions (Repeat and Expansion of Bairey et al.)"
author: "Chris Terry"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
editor_options: 
  chunk_output_type: console
---

# Introduction 

This script documents the analysis to first repeat the analysis of Bairey et al. 2016 (Bairey, E., Kelsic, E. & Kishony, R. High-order species interactions shape ecosystem diversity. Nat Commun 7, 12285 (2016). https://doi.org/10.1038/ncomms12285) and then extend it with structured HOIs. 


```{r message=FALSE, warning=FALSE}
#library(pracma)
library(deSolve)
sapply(paste0('Scripts/',list.files('Scripts/')), source)->tmp
require(igraph)
require(MASS)
require(tidyverse)
require(pracma)
require(moments)
require(rmutil)
require(mvtnorm)
require(mgcv)

```


# ODE Models 

To minimise operations each combination of orders of interactions is made as a separate model, but all follow the same format.

All models are LV rather than replicator equations, so technically match SI 5 of Bairey et al, not the main text figures. 

```{R}
PairDyn = function(t, state, parameters ){
  if( any(is.na(state))|any(state<0.000001/N)){return(list(rep(0,N))) } ## if a sure gonner, stop calculating immediately.
  x_i <-  - state## Self Reg 
  Pairwise<-    A_mat%*%state## pairwise effects
  return( list(state*(r+ x_i + Pairwise )))
}
ThreeDyn = function(t, state, parameters ){
  
  x_i <-  - state ## Self Reg 
  ## Three way effects
  XjXk = outer(state,state)
  ThreeWay<- rep(NA, N) 
  for( i in 1:N){
    ThreeWay[i] <- sum(B_mat[i,,]*XjXk)
  }
  return(list(state*(r+ x_i +ThreeWay)))
}
FourDyn = function(t, state, parameters ){
  
  if( any(is.na(state))|any(state<0.000001/N)){return(list(rep(0,N))) }  
  ## Self Reg 
  x_i <-  - state
  ## four way effects i,j,k,l
  Fourway<- rep(NA, N) ## vector of effects on each i
  for( i in 1:N){
    XkXl = outer(state,state)
    XjXkXl =  sweep(array(XkXl, dim = c(N,N,N) ),3,state,FUN="*")
    Fourway[i] <- sum(C_mat[i,,,]*XjXkXl)
  }
  return( list(state*(r+ x_i  + Fourway)))
}
TwoAndThreeDyn = function(t, state, parameters ){
  if( any(is.na(state))|any(state<0.000001/N)){return(list(rep(0,N))) } 
  x_i <-  - state ## Self Reg
  Pairwise<-    A_mat%*%state  ## Pairwise
  ## Three way effects
  XjXk = outer(state,state)
  ThreeWay<- rep(NA, N) 
  for( i in 1:N){
    ThreeWay[i] <- sum(B_mat[i,,]*XjXk)
  }
  return(list(state*(r+ x_i + Pairwise+ ThreeWay)))
}

PairAndFourDyn = function(t, state, parameters ){
  if( any(is.na(state))|any(state<0.000001/N)){return(list(rep(0,N))) } 
  x_i <-  - state  ## Self Reg
  Pairwise<-    A_mat%*%state ## Pairwise
  ## four way effects i,j,k,l
  Fourway<- rep(NA, N) ## vector of effects on each i
  for( i in 1:N){
    XkXl = outer(state,state)
    XjXkXl =  sweep(array(XkXl, dim = c(N,N,N) ),3,state,FUN="*")
    Fourway[i] <- sum(C_mat[i,,,]*XjXkXl)
  }
  return( list(state*(r+ x_i  +Pairwise+  Fourway)))
}

```

# Process of finding critical alpha's

For each number of species (N), calculate the critical strength of an interaction that leads to loss of feasibility. 

Feasibility is defined as >10% of random draws having at least one extinction. Extinctions are defined as a species being below 0.0001/N at the end point. 

For computational efficiency, integration is interrupted if a species falls below 0.000001/N, Convergence is tested by a comparison of t = 700 with t = 1000. If the entropy has stabilised the run is stopped, if not, it is run to 10000 time steps.   Plots are made to check up on whether this is enough. 

Strengths are tested from high values and successively reduced on a logarithmic scale. After 10% of runs fail, the strength is reduced. 


## Example Simplified loop

This loop has additional comments and the  diagnostics and some error catchers removed for clarity. Others follow the same structure. 

```{r eval = FALSE}
N_timesteps = 10  ## number of timesteps to record from the ode. (by having fewer, the integrator can make bigger leaps)
TimeStepsInt = seq( 1, 1000, length = N_timesteps)  ## time steps to use in first integration.

N_reps = 50   # number of communities to build at each strength level
Str_seq = rev(10^seq(from= -3, to = -1, by = 0.1 ))  # sequence of strengths of interaction to test (different for each) 
N_Seq =  seq(5, 25, by = 5)          # sequence of species numbers to test. 

N_df_alpha = data.frame(N =  N_Seq,  CritInt = NA) ## data.frame to save critical values in. 

for(N_i  in 1:length(N_Seq)){ ## cycling through species numbers
  N = N_df_alpha$N[N_i]
  for ( alpha  in  Str_seq ){    ## cycling down through critical strengths
    Feasible_vec = rep(NA, N_reps)  ## make vector to record feasibility 
    for( rep in 1:N_reps){           ## cycling through replicates
      if(sum(Feasible_vec==FALSE, na.rm=TRUE)> 0.1*N_reps){       ## If already know enough are not feasible at this strength, don't run more integration
        Feasible_vec[rep]<-FALSE
      }else{
        set.seed(rep)  # setting the seed here means the distribution of interactions matches across strengths
        r = 1/N    #Self effects 
        A_mat = matrix(rnorm(N*N), nrow = N, ncol = N) *sqrt(alpha)  ## generate interaction matrices
        Traj <-  ode(f = PairDyn, times = TimeStepsInt,             ## run ODE 
                     y = rep(1/N, N),                                 # start values = 1/N
                     method = 'ode45', parms = NULL)
        Endpoints<- Traj[,-1][N_timesteps,] # save endpoint
        SeventyPerc<- Traj[,-1][round(N_timesteps*0.7),] # save data from 70% of the way through time series
        
        if(any(is.na(Endpoints)) | any(Endpoints<=0) |any(SeventyPerc<=0)  ){  ## if any extinct, declare not feasible
          Feasible_vec[rep]<- FALSE 
        }else{
          Ent1=sum( Endpoints*log(Endpoints)) ## Calc two values of entropy to detect if still getting movement
          Ent2=sum( SeventyPerc*log(SeventyPerc))
          if( log10(abs(Ent1-Ent2)) > -5){ ## if not converged by this point, run for longer
            Traj <-  ode(f = PairDyn, times = seq(1000, 10000, length = N_timesteps),
                         y = Endpoints, method = 'ode45', parms = NULL)
            Endpoints<- Traj[,-1][N_timesteps,] ## save extended endpoints
          }
          Feasible_vec[rep]<- (all(Endpoints>0.0001/N  ) &   all(Endpoints< 100*N  ))  ## also fail if any species is very large
          if(is.na(Feasible_vec[rep])){Feasible_vec[rep]<- FALSE }  ## error catching 
        }
      }
    }
    frac_feas = sum(Feasible_vec, na.rm=TRUE)/N_reps  # calculate fesiblity
    if( frac_feas > 0.9){   # if feasibility >90% identified, 
      N_df_alpha$CritInt[N_i] = alpha  ## then declare the current alpha the critical value 
      break()           ## don't test even weaker strengths. 
    }
  }
}

```

# Finding Critical Values

### Consistent Values 

```{r}
N_timesteps = 10
TimeStepsInt = seq( 1, 1000, length = N_timesteps)
N_reps = 50
N_Seq =  seq(10, 30, by = 5)

PLOT = FALSE ## don't make diagnostic plots anymore

```


## Finding Critical alpha value for pairwise

```{r  eval = FALSE}
Str_seq = rev(10^seq(from= -3, to = -1, by = 0.05 ))
N_df_alpha = data.frame(N =  N_Seq,  CritInt = NA)

for(N_i  in 1:length(N_Seq)){
  N = N_df_alpha$N[N_i]
  print(paste('N=', N))
  for ( alpha  in  Str_seq ){
    print(paste('alpha=', log10(alpha)))
    Feasible_vec = rep(NA, N_reps)
    for( rep in 1:N_reps){
      if(sum(Feasible_vec==FALSE, na.rm=TRUE)> 0.1*N_reps){       ## if already had more than 10% of total fails, don't bother running more replicates at this level
        Feasible_vec[rep]<-FALSE
        cat('E')
      }else{
        set.seed(rep)
        r = 1/N    #Self effects 
        A_mat = matrix(rnorm(N*N), nrow = N, ncol = N) *sqrt(alpha)
        Traj <-  ode(f = PairDyn, times = TimeStepsInt,
                     y = rep(1/N, N), method = 'ode45', parms = NULL)
        Endpoints<- Traj[,-1][N_timesteps,] # save endpoint
        SeventyPerc<- Traj[,-1][round(N_timesteps*0.7),] # save data from 70% of the way through time series
        
        if(any(is.na(Endpoints)) | any(Endpoints<=0) |any(SeventyPerc<=0)  ){
          Feasible_vec[rep]<- FALSE 
        }else{
          ## Calc two values of entropy to detect if still getting movement
          Ent1=sum( Endpoints*log(Endpoints))
          Ent2=sum( SeventyPerc*log(SeventyPerc))
          if( log10(abs(Ent1-Ent2)) > -5){ ## if not converged by this point, run for longer
            cat('_X')
            if(PLOT){ matplot(log10(Traj[,-1]), type = 'l'); title( paste('Running Extra Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
            Traj <-  ode(f = PairDyn, times = seq(1000, 10000, length = N_timesteps),
                         y = Endpoints, method = 'ode45', parms = NULL)
            Endpoints<- Traj[,-1][N_timesteps,] 
          }
          Feasible_vec[rep]<- (all(Endpoints>0.0001/N  ) &   all(Endpoints< 100*N  ))  
          if(is.na(Feasible_vec[rep])){Feasible_vec[rep]<- FALSE }
        }
        cat(Feasible_vec[rep]*1)
        if(PLOT){   matplot(log10(Traj[,-1]), type = 'l'); title( paste('Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
      }
    }
    
    frac_feas = sum(Feasible_vec, na.rm=TRUE)/N_reps
    print(frac_feas)
    if( frac_feas > 0.9){ ## reversing 
      N_df_alpha$CritInt[N_i] = alpha
      break()
    }
  }
}

N_df_alpha$LogAlphaCrit = log10(N_df_alpha$CritInt)
write_csv(N_df_alpha,   'Generated Data/CriticalValues/N_df_alpha.csv')
plot(x=N_df_alpha$N, y = log10(N_df_alpha$CritInt))

```

## Finding Critical beta value for three-way

```{r eval = FALSE}
Str_seq = rev(10^seq(from= -1.5, to = 0, by = 0.05 ))
N_df_beta = data.frame(N =  N_Seq, CritInt = NA)

for(N_i  in 1:length(N_Seq)){
  N = N_df_beta$N[N_i]
  print(paste('N=', N))
  for ( beta  in  Str_seq ){
    print(paste('beta=', log10(beta)))
    Feasible_vec = rep(NA, N_reps)
    for( rep in 1:N_reps){
      if(sum(Feasible_vec==FALSE, na.rm=TRUE)> 0.1*N_reps){       ## if already had more than 10% of total fails, don't bother running more replicates at this level
        Feasible_vec[rep]<-FALSE
        cat('E')
      }else{
        set.seed(rep)
        r = 1/N    #Self effects 
        B_mat = array(rnorm(N^3), dim = c(N,N,N))* sqrt(beta)## Threeway
        Traj <-  ode(f = ThreeDyn, times = TimeStepsInt,
                     y = rep(1/N, N), method = 'ode45', parms = NULL)
        Endpoints<- Traj[,-1][N_timesteps,] # save endpoint
        SeventyPerc<- Traj[,-1][round(N_timesteps*0.7),] # save data from 70% of the way through time series
        
        if(any(is.na(Endpoints)) | any(Endpoints<=0) |any(SeventyPerc<=0) ){
          Feasible_vec[rep]<- FALSE 
        }else{
          ## Calc two values of entropy to detect if still getting movement
          Ent1=sum( Endpoints*log(Endpoints))
          Ent2=sum( SeventyPerc*log(SeventyPerc))
          if( log10(abs(Ent1-Ent2)) > -5){ ## if not converged by this point, run for longer
            cat('_X')
            if(PLOT){ matplot(log10(Traj[,-1]), type = 'l'); title( paste('Running Extra Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
            Traj <-  ode(f = ThreeDyn, times = seq(1000, 10000, length = N_timesteps),
                         y = Endpoints, method = 'ode45', parms = NULL)
            Endpoints<- Traj[,-1][N_timesteps,] 
          }
          Feasible_vec[rep]<- (all(Endpoints>0.0001/N  ) &   all(Endpoints< 100*N  ))  
          if(is.na(Feasible_vec[rep])){Feasible_vec[rep]<- FALSE }
        }
        cat(Feasible_vec[rep]*1)
        if(PLOT){   matplot(log10(Traj[,-1]), type = 'l'); title( paste('Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
      }
    }
    
    frac_feas = sum(Feasible_vec, na.rm=TRUE)/N_reps
    print(frac_feas)
    if( frac_feas > 0.9){ ## reversing 
      N_df_beta$CritInt[N_i] = beta
      break()
    }
  }
}

N_df_beta$LogBetaCrit = log10(N_df_beta$CritInt)

write_csv(N_df_beta,  'Generated Data/CriticalValues/N_df_beta.csv')

plot(x=N_df_beta$N, y = log10(N_df_beta$CritInt))

```

## Finding critical gamma value (Four way)  


```{r eval = FALSE}
Str_seq = rev(10^seq(from= -1.5, to = 0.6, by = 0.05 ))
N_df_gamma = data.frame(N =  N_Seq,  CritInt = NA)

for(N_i  in 1:length(N_Seq)){
  N = N_df_gamma$N[N_i]
  print(paste('N=', N))
  for ( gamma  in  Str_seq ){
    print(paste('gamma=', log10(gamma)))
    Feasible_vec = rep(NA, N_reps)
    for( rep in 1:N_reps){
      if(sum(Feasible_vec==FALSE, na.rm=TRUE)> 0.1*N_reps){
        ## if already had more than 10% of total fails, don't bother running more
        Feasible_vec[rep]<-FALSE
        cat('E')
      }else{
        
        set.seed(rep)
        #Self effects 
        r = 1/N
        ## Fourway effects
        C_mat = array(rnorm(N^4), dim = c(N,N,N,N))* sqrt(gamma)
        Traj <-  ode(f = FourDyn, times = TimeStepsInt,
                     y = rep(1/N, N), method = 'ode45', parms = NULL)
        
        Endpoints<- Traj[,-1][N_timesteps,]
        SeventyPerc<- Traj[,-1][round(N_timesteps*0.7),]
        
        if(any(is.na(Endpoints) | any(Endpoints<=0) |any(SeventyPerc<=0) ) ){
          Feasible_vec[rep]<- FALSE 
        }else{
          ## Calc two values of entropy
          Ent1=sum( Endpoints*log(Endpoints))
          Ent2=sum( SeventyPerc*log(SeventyPerc))
          
          if( log10(abs(Ent1-Ent2)) > -5){ ## if not converged by this point, run for longer
            cat('_X')
             if(PLOT){  matplot(log10(Traj[,-1]), type = 'l'); title( paste('Running Extra Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
            Traj <-  ode(f = FourDyn, times = seq(1000, 10000, length = N_timesteps),
                         y = Endpoints, method = 'ode45', parms = NULL)
            Endpoints<- Traj[,-1][N_timesteps,] 
          }
          Feasible_vec[rep]<- (all(Endpoints>0.0001/N  ) &   all(Endpoints< 100*N  ))  
          if(is.na(Feasible_vec[rep])){Feasible_vec[rep]<- FALSE }
        }
        cat(Feasible_vec[rep]*1)
         if(PLOT){  matplot(log10(Traj[,-1]), type = 'l'); title( paste('Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
      }
    }
    
    frac_feas = sum(Feasible_vec, na.rm=TRUE)/N_reps
    print(frac_feas)
    if( frac_feas > 0.9){ ## reversing 
      N_df_gamma$CritInt[N_i] = gamma
      break()
    }
  }
}

plot(x=N_df_gamma$N, y = log10(N_df_gamma$CritInt))
N_df_gamma$LogGammaCrit = log10(N_df_gamma$CritInt)
write_csv(N_df_gamma,  'Generated Data/CriticalValues/N_df_gamma.csv')
```

# Repeating Bairey with constant trophic underlay and random 3 and 4 way:

## Three way on top

`A` mat added at a level that allows feasibility at that population size. 

```{r}
Build_A <- function( rep, N){return(BuildWeb(ID =rep, N,  connectance = 0.5,
                                             mux = -1,  muy = 1,
                                             sigmax = 1,  sigmay = 1, 
                                             rhoxy = -2/3))}
```

## Random 3way

```{R eval = FALSE}
Str_seq = rev(10^seq(from= -2, to = 0, by = 0.05))
alpha = 0.001 ## Strength of underlying trophic network
N_df_beta_RAND_ONTROPHIC = data.frame(N =  N_Seq, CritInt = NA)


for(N_i  in 1:length(N_Seq)){
  N = N_df_beta_RAND_ONTROPHIC$N[N_i]
  print(paste('N=', N))
  for ( beta  in  Str_seq ){
    print(paste('beta=', log10(beta)))
    Feasible_vec = rep(NA, N_reps)
    for( rep in 1:N_reps){
      if(sum(Feasible_vec==FALSE, na.rm=TRUE)> 0.1*N_reps){       ## if already had more than 10% of total fails, don't bother running more replicates at this level
        Feasible_vec[rep]<-FALSE
        cat('E')
      }else{
        set.seed(rep)
        r = 1/N    #Self effects 
        ### defining underlying tropic web#####################
        A_mat_unscaled = Build_A(rep, N)
        A_mat <- A_mat_unscaled/ mean(abs(A_mat_unscaled)) *sqrt(alpha) / sqrt(N)
        #        image(A_mat, main = 'Trophic network')
        ################################
        set.seed(rep)
        
        B_mat = array(rnorm(N^3), dim = c(N,N,N))* sqrt(beta)## Threeway
        Traj <-  ode(f = TwoAndThreeDyn, times = TimeStepsInt,
                     y = rep(1/N, N), method = 'ode45', parms = NULL)
        Endpoints<- Traj[,-1][N_timesteps,] # save endpoint
        SeventyPerc<- Traj[,-1][round(N_timesteps*0.7),] # save data from 70% of the way through time series
        
        if(any(is.na(Endpoints)) | any(Endpoints<=0) |any(SeventyPerc<=0) ){
          Feasible_vec[rep]<- FALSE 
        }else{
          ## Calc two values of entropy to detect if still getting movement
          Ent1=sum( Endpoints*log(Endpoints))
          Ent2=sum( SeventyPerc*log(SeventyPerc))
          if( log10(abs(Ent1-Ent2)) > -5){ ## if not converged by this point, run for longer
            cat('_X')
            if(PLOT){ matplot(log10(Traj[,-1]), type = 'l'); title( paste('Running Extra Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
            Traj <-  ode(f = TwoAndThreeDyn, times = seq(1000, 10000, length = N_timesteps),
                         y = Endpoints, method = 'ode45', parms = NULL)
            Endpoints<- Traj[,-1][N_timesteps,] 
          }
          Feasible_vec[rep]<- (all(Endpoints>0.0001/N  ) &   all(Endpoints< 100*N  ))  
          if(is.na(Feasible_vec[rep])){Feasible_vec[rep]<- FALSE }
        }
        cat(Feasible_vec[rep]*1)
        if(PLOT){   matplot(log10(Traj[,-1]), type = 'l'); title( paste('Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
      }
    }
    
    frac_feas = sum(Feasible_vec, na.rm=TRUE)/N_reps
    print(frac_feas)
    if( frac_feas > 0.9){ ## reversing 
      N_df_beta_RAND_ONTROPHIC$CritInt[N_i] = beta
      break()
    }
  }
}

N_df_beta_RAND_ONTROPHIC$LogBetaCrit = log10(N_df_beta_RAND_ONTROPHIC$CritInt) ## Critical value stays constant at ~1.2 with random HOIs. 
write_csv(N_df_beta_RAND_ONTROPHIC, 
          'Generated Data/CriticalValues/N_df_beta_RAND_ONTROPHIC.csv')
```

## Structured 3way

```{R eval = FALSE}
Str_seq  =rev(10^seq(from= -3, to = -1, by = 0.05  ))

alpha = 0.001
TIM_Dens =0.1        ## How many interactions does each species modify? [[proportion of cube non-zero]]
N_df_beta_STRUC_ONTROPHIC = data.frame(N =  N_Seq,  CritInt = NA)

for(N_i  in 1:length(N_Seq)){
  N = N_df_beta_STRUC_ONTROPHIC$N[N_i]
  print(paste('N=', N))
  for ( beta  in  Str_seq ){
    print(paste('beta=', log10(beta)))
    Feasible_vec = rep(NA, N_reps)
    for( rep in 1:N_reps){
      if(sum(Feasible_vec==FALSE, na.rm=TRUE)> 0.1*N_reps){       ## if already had more than 10% of total fails, don't bother running more replicates at this level
        Feasible_vec[rep]<-FALSE
        cat('E')
      }else{
        set.seed(rep)
        r = 1/N    #Self effects 
        ### defining underlying tropic web#####################
        A_mat_unscaled = Build_A(rep, N)
        A_mat <- A_mat_unscaled/ mean(abs(A_mat_unscaled)) *sqrt(alpha) / sqrt(N)
        #        image(A_mat, main = 'Trophic network')
        ################################
        set.seed(rep)
        ### Build structured three way
        B_mat_unscaled <- Build_Barray_Switching(A_mat, TIM_Dens = TIM_Dens*N*N)  ## N^2 here maintains connectance of HOI cube at approximately TIM_dens 
        B_mat_scaled_var1 = B_mat_unscaled / sd( B_mat_unscaled) 
        B_mat = B_mat_scaled_var1 * sqrt(beta)
        #  print(paste0('TrophicConn=',mean(A_mat!=0),'TIM_CONN=', mean(B_mat!=0)))
        Traj <-  ode(f = TwoAndThreeDyn, times = TimeStepsInt,
                     y = rep(1/N, N), method = 'ode45', parms = NULL)
        Endpoints<- Traj[,-1][N_timesteps,] # save endpoint
        SeventyPerc<- Traj[,-1][round(N_timesteps*0.7),] # save data from 70% of the way through time series
        
        if(any(is.na(Endpoints)) | any(Endpoints<=0) |any(SeventyPerc<=0) ){
          Feasible_vec[rep]<- FALSE 
        }else{
          ## Calc two values of entropy to detect if still getting movement
          Ent1=sum( Endpoints*log(Endpoints))
          Ent2=sum( SeventyPerc*log(SeventyPerc))
          if( log10(abs(Ent1-Ent2)) > -5){ ## if not converged by this point, run for longer
            cat('_X')
            if(PLOT){ matplot(log10(Traj[,-1]), type = 'l'); title( paste('Running Extra Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
            Traj <-  ode(f = TwoAndThreeDyn, times = seq(1000, 10000, length = N_timesteps),
                         y = Endpoints, method = 'ode45', parms = NULL)
            Endpoints<- Traj[,-1][N_timesteps,] 
          }
          Feasible_vec[rep]<- (all(Endpoints>0.0001/N  ) &   all(Endpoints< 100*N  ))  
          if(is.na(Feasible_vec[rep])){Feasible_vec[rep]<- FALSE }
        }
        cat(Feasible_vec[rep]*1)
        if(PLOT){   matplot(log10(Traj[,-1]), type = 'l'); title( paste('Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
      }
    }
    
    frac_feas = sum(Feasible_vec, na.rm=TRUE)/N_reps
    print(frac_feas)
    if( frac_feas > 0.9){ ## reversing 
      N_df_beta_STRUC_ONTROPHIC$CritInt[N_i] = beta
      break()
    }
  }
}

N_df_beta_STRUC_ONTROPHIC$LogBetaCrit = log10(N_df_beta_STRUC_ONTROPHIC$CritInt) 
write_csv(N_df_beta_STRUC_ONTROPHIC, 
          'Generated Data/CriticalValues/N_df_beta_STRUC_ONTROPHIC.csv')
## With structure, get more destabilising impact of HOIs
```

## 4 way 

### Random 4way overlaid over trophic

```{r eval = FALSE}
alpha = 0.001
Str_seq =  rev(10^seq(from= -2, to = 1, by = 0.05  ))
N_df_gamma_RANDOM_ONTROPHIC = data.frame(N =  N_Seq, CritInt = NA)

for(N_i  in 1:length(N_Seq)){
  N = N_df_gamma_RANDOM_ONTROPHIC$N[N_i]
  print(paste('N=', N))
  #Self effects 
  r = 1/N
  for ( gamma  in  Str_seq ){
    print(paste('gamma=', log10(gamma)))
    Feasible_vec = rep(NA, N_reps)
    for( rep in 1:N_reps){
      if(sum(Feasible_vec==FALSE, na.rm=TRUE)> 0.1*N_reps){        ## if already had more than 10% of total fails, don't bother running more
        Feasible_vec[rep]<-FALSE;cat('E')
      }else{
        set.seed(rep)
        # Trophic Interactions
        A_mat_unscaled = Build_A(rep, N)
        A_mat <- A_mat_unscaled/ mean(abs(A_mat_unscaled)) *sqrt(alpha) / sqrt(N)
        ################################
        set.seed(rep)
        
        ## Random Fourway effects
        C_mat = array(rnorm(N^4), dim = c(N,N,N,N))* sqrt(gamma)
        Traj <-  ode(f = PairAndFourDyn, times = TimeStepsInt,
                     y = rep(1/N, N), method = 'ode45', parms = NULL)
        
        Endpoints<- Traj[,-1][N_timesteps,]
        SeventyPerc<- Traj[,-1][round(N_timesteps*0.7),]
        
        if(any(is.na(Endpoints) | any(Endpoints<=0) |any(SeventyPerc<=0) ) ){
          Feasible_vec[rep]<- FALSE 
        }else{
          ## Calc two values of entropy
          Ent1=sum( Endpoints*log(Endpoints))
          Ent2=sum( SeventyPerc*log(SeventyPerc))
          
          if( log10(abs(Ent1-Ent2)) > -5){ ## if not converged by this point, run for longer
            cat('_X')
            if(PLOT){    matplot(log10(Traj[,-1]), type = 'l'); title( paste('Running Extra Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
            Traj <-  ode(f = PairAndFourDyn, times = seq(1000, 10000, length = N_timesteps),
                         y = Endpoints, method = 'ode45', parms = NULL)
            Endpoints<- Traj[,-1][N_timesteps,] 
          }
          Feasible_vec[rep]<- (all(Endpoints>0.0001/N  ) &   all(Endpoints< 100*N  ))  
          if(is.na(Feasible_vec[rep])){Feasible_vec[rep]<- FALSE }
        }
        cat(Feasible_vec[rep]*1)
       if(PLOT){     matplot(log10(Traj[,-1]), type = 'l'); title( paste('Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
      }
    }
    
    frac_feas = sum(Feasible_vec, na.rm=TRUE)/N_reps
    print(frac_feas)
    if( frac_feas > 0.9){ ## reversing 
      N_df_gamma_RANDOM_ONTROPHIC$CritInt[N_i] = gamma
      break()
    }
  }
}

N_df_gamma_RANDOM_ONTROPHIC$LogGammaCrit = log10(N_df_gamma_RANDOM_ONTROPHIC$CritInt) 

write_csv(N_df_gamma_RANDOM_ONTROPHIC, 
          'Generated Data/CriticalValues/N_df_gamma_RANDOM_ONTROPHIC.csv')

```

### Structured 4way overlaid over trophic

Not really any expectations of 4-way, but assume of [i,j,k,l] array, where i and j are prey and predator, k is a modifier, l is double modifier. 

Define [i,j,k, ] as per 3-way structure, then define l axis randomly 

```{r eval = FALSE}
alpha = 0.001
Str_seq =   rev(10^seq(from= -2, to = 0.5, by = 0.05  ))
N_df_gamma_STRUCT_OnTROPHIC = data.frame(N =  N_Seq, CritInt = NA)

for(N_i  in 1:length(N_Seq)){
  N = N_df_gamma_STRUCT_OnTROPHIC$N[N_i]
  print(paste('N=', N))
  #Self effects 
  r = 1/N
  for ( gamma  in  Str_seq ){
    print(paste('gamma=', log10(gamma)))
    Feasible_vec = rep(NA, N_reps)
    for( rep in 1:N_reps){
      if(sum(Feasible_vec==FALSE, na.rm=TRUE)> 0.1*N_reps){        ## if already had more than 10% of total fails, don't bother running more
        Feasible_vec[rep]<-FALSE;cat('E')
      }else{
        set.seed(rep)
        # Trophic Interactions
        A_mat_unscaled = Build_A(rep, N)
        A_mat <- A_mat_unscaled/ mean(abs(A_mat_unscaled)) *sqrt(alpha) / sqrt(N)
        ################################
        set.seed(rep)
        
        ## Random Fourway effects
        ## Find potential three way effects based on trophic network
        Threeway_array_binary <- Build_Barray_Switching(A_mat, TIM_Dens = TIM_Dens*N*N)  ## N^2 here maintains connectance of HOI cube at approximately TIM_dens 
        C_arr_binary <- array(NA, dim = c(N,N,N,N))
        for( l in 1:N){
          C_arr_binary[,,,l] <- Threeway_array_binary * rbinom(n=N^3, size =1,prob = 0.1) ## 10% of each TIM being modified
        }
        C_mat = C_arr_binary/ sd(C_arr_binary) * sqrt(gamma)
        

        Traj <-  ode(f = PairAndFourDyn, times = TimeStepsInt,
                     y = rep(1/N, N), method = 'ode45', parms = NULL)
        
        Endpoints<- Traj[,-1][N_timesteps,]
        SeventyPerc<- Traj[,-1][round(N_timesteps*0.7),]
        
        if(any(is.na(Endpoints) | any(Endpoints<=0) |any(SeventyPerc<=0) ) ){
          Feasible_vec[rep]<- FALSE 
        }else{
          ## Calc two values of entropy
          Ent1=sum( Endpoints*log(Endpoints))
          Ent2=sum( SeventyPerc*log(SeventyPerc))
          
          if( log10(abs(Ent1-Ent2)) > -5){ ## if not converged by this point, run for longer
            cat('_X')
           if(PLOT){     matplot(log10(Traj[,-1]), type = 'l'); title( paste('Running Extra Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
            Traj <-  ode(f = PairAndFourDyn, times = seq(1000, 10000, length = N_timesteps),
                         y = Endpoints, method = 'ode45', parms = NULL)
            Endpoints<- Traj[,-1][N_timesteps,] 
          }
          Feasible_vec[rep]<- (all(Endpoints>0.0001/N  ) &   all(Endpoints< 100*N  ))  
          if(is.na(Feasible_vec[rep])){Feasible_vec[rep]<- FALSE }
        }
        cat(Feasible_vec[rep]*1)
      if(PLOT){matplot(log10(Traj[,-1]), type = 'l'); title( paste('Rep:', rep, Feasible_vec[rep]));Sys.sleep(0.1)}
      }
    }
    
    frac_feas = sum(Feasible_vec, na.rm=TRUE)/N_reps
    print(frac_feas)
    if( frac_feas > 0.9){ ## reversing 
      N_df_gamma_STRUCT_OnTROPHIC$CritInt[N_i] = gamma
      break()
    }
  }
}

N_df_gamma_STRUCT_OnTROPHIC$LogGammaCrit = log10(N_df_gamma_STRUCT_OnTROPHIC$CritInt) 

write_csv(N_df_gamma_STRUCT_OnTROPHIC, 
          'Generated Data/CriticalValues/N_df_gamma_STRUCT_OnTROPHIC.csv')

plot(x=N_df_gamma_STRUCT_OnTROPHIC$N, y = N_df_gamma_STRUCT_OnTROPHIC$LogGammaCrit)
```

# Plotting all 

```{r message=FALSE, warning=FALSE}
fp <- 'Generated Data/CriticalValues/N_df_'

bind_rows(
  read_csv(paste0(fp,'alpha.csv')) %>% mutate(Exp='a) Random HOIs', Order='Two-way'),
  read_csv(paste0(fp,'beta.csv')) %>% mutate(Exp='a) Random HOIs', Order='Three-way'),
  read_csv(paste0(fp,'gamma.csv')) %>% mutate(Exp='a) Random HOIs', Order='Four-way'),
  read_csv(paste0(fp,'beta_RAND_ONTROPHIC.csv')) %>% mutate(Exp='b) Trophic Structure and Random HOIs', Order='Three-way'),
  read_csv(paste0(fp,'gamma_RANDOM_ONTROPHIC.csv')) %>% mutate(Exp='b) Trophic Structure and Random HOIs', Order='Four-way'),
  read_csv(paste0(fp,'beta_STRUC_ONTROPHIC.csv')) %>% mutate(Exp='c) Trophic Structure and Structured HOIs', Order='Three-way'),
  read_csv(paste0(fp,'gamma_STRUCT_OnTROPHIC.csv')) %>% mutate(Exp='c) Trophic Structure and Structured HOIs', Order='Four-way')
) -> All_Criticals

All_Criticals %>%
  ggplot(aes( x = N, y =  CritInt, col=Order))+
    geom_smooth(method = 'lm', se = FALSE, alpha= 0.5)+
geom_point( )+
  scale_y_log10()+
  scale_color_viridis_d(name = 'Order of\ninteractions')+
  theme_bw()+
  facet_wrap(~ Exp)+
  ylab('Critical Strength\nof Interactions')+
  xlab('Number of species, N')


```

# Session Info

```{r}
sessionInfo()
```



